---
title: "Changing established atttiudes via the IAT: Experiment 3"
author: "Ian Hussey"
output:
  html_document:
    highlight: haddock
    toc: yes
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# dependencies
library(tidyverse)
library(psych)
library(weights)  # for rd(), a round() alternative
library(plotrix)  # for std.error
library(lme4)
library(effects)
library(knitr)
library(sjPlot)
library(kableExtra)

# comment out if knitting to pdf
options(knitr.table.format = "html")

# knitr options
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

# rounds all numerics in a df
round_df <- function(df, digits) {
  nums <- vapply(df, is.numeric, FUN.VALUE = logical(1))
  df[,nums] <- round(df[,nums], digits = digits)
  (df)
}
```

# Descriptive statistics

## Full sample

### Age

```{r, message=FALSE, warning=FALSE, echo = FALSE}
# get data
data_df <- 
  read.csv("data/processed/processed summary data - wide format.csv") %>%
  dplyr::mutate(gender = as.factor(gender))

#colnames(data_df)

data_df %>% 
  dplyr::select(age) %>%
  psych::describe(fast = TRUE,  # subset of descriptive stats
                  ranges = FALSE,
                  trim = 0) %>%
  dplyr::select(-vars, -se) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### Gender

```{r, message=FALSE, warning=FALSE, echo = FALSE}
data_df %>% 
  dplyr::count(gender) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

## Analytic sample

I.e., after applying inclusion/exclusion criteria.

### Conditions

```{r, message=FALSE, warning=FALSE, echo=FALSE}
passers_df <- 
  data_df %>%
  filter(IAT_exclude_based_on_fast_trials == FALSE &
           shooter_exclude_based_on_fast_trials == FALSE)

passers_df %>% 
  dplyr::count(IAT_condition) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### Performance on the IAT and shooter task

```{r, message=FALSE, warning=FALSE, echo=FALSE}
passers_df %>% 
  dplyr::select(IAT_mean_RT,
                IAT_perc_acc,
                shooter_mean_RT, 
                shooter_perc_acc) %>%
  psych::describe(fast = TRUE,  # subset of descriptive stats
                  ranges = FALSE,
                  trim = 0) %>%
  tibble::rownames_to_column() %>%
  dplyr::select(-vars, -se) %>%
  dplyr::mutate(mean = round(mean, 2),
         sd = round(sd, 2)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

# Hypothesis test 1: Differences in IAT effects between contrast pair conditions

H1-H3: Completing an IAT serves to train attitudes as well as test them. Participants who complete a race IAT will influence implicit biases towards black people on the on the shooter task relative to participants who completed a control (flowers-insects) IAT.

H1 refers to the RT analysis: main effect for condition or interaction effect between condition and trial type.

## Outlier removal

Exclude data points > 2.5 SD from mean.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# exclude participants who had more than 10% of trials in less than 300ms, and make variables factors
shooter_data <- 
  read.csv("data/processed/processed shooter task data - long format.csv") %>%
  dplyr::mutate(participant = as.factor(participant),
         trial_type = as.factor(trial_type),
         IAT_condition = as.factor(IAT_condition)) %>%
  filter(IAT_exclude_based_on_fast_trials == FALSE &
           shooter_exclude_based_on_fast_trials == FALSE)

# exclude outlier rts
shooter_data_outliers_removed <- shooter_data %>%
  schoRsch::outlier(dv = "rt",
                    todo="elim", 
                    upper.z = 2.5, 
                    lower.z = -2.5)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
# Check that variables that should be factors are indeed factors
sapply(shooter_data_outliers_removed, class)
```

### Plot distribution

Distribution of shooter RTs after trimming outliers.

```{r, echo=FALSE, fig.height=3, fig.width=5, message=FALSE, warning=FALSE}
ggplot(data = shooter_data_outliers_removed) + 
  geom_density(aes(x = rt)) +
  theme_classic()
```

## Linear mixed model

### Model

```{r, message=FALSE, warning=FALSE}
model_1 <- lmer(rt ~ trial_type * IAT_condition + modern_racism_scale_total + (1 | participant), 
                contrasts = list(trial_type = "contr.sum", IAT_condition = "contr.sum"),  
                data = shooter_data_outliers_removed)
```

- Hypothesis refers to the main effect for IAT condition or the IAT condition*trial type interaction effect.
- Confidence intervals via Wald estimation.

### Summary

```{r, echo=FALSE, message=FALSE, warning=FALSE}
sjt.lmer(model_1, 
         emph.p = FALSE, 
         p.kr = FALSE,
         show.std = TRUE)
```

### Plot effects

```{r, echo=FALSE, fig.height=3, fig.width=5, message=FALSE, warning=FALSE}
sjp.lmer(model_1, 
         p.kr = FALSE,
         type = "fe")
```

### Table predictions

```{r, echo=FALSE, fig.height=3, fig.width=5, message=FALSE, warning=FALSE}
model_1_predicted_effects <- 
  as.data.frame(effect("trial_type:IAT_condition", model_1)) %>%
  dplyr::rename(`IAT condition` = IAT_condition,
         `Trial type` = trial_type)

model_1_predicted_effects %>% 
  round_df(2) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```
  
### Plot predictions

```{r, echo=FALSE, fig.height=3, fig.width=5, message=FALSE, warning=FALSE}
ggplot(data = model_1_predicted_effects,
       aes(x = `Trial type`, y = fit, colour = `IAT condition`)) +
  geom_pointrange(aes(ymax = upper,
                      ymin = lower),
                  position = position_dodge(width = .1)) +
  geom_line(aes(group = `IAT condition`),
            position = position_dodge(width = .1)) +
  ylab("RT") +
  xlab("Trial type") +
  scale_colour_grey() +
  theme_classic()
  # theme(legend.justification=c(0.5,1), legend.position=c(0.5,1),
  #       legend.key = element_rect(colour = "transparent"))
```

# Hypothesis test 2: differences in sensitivity (d')

H2 refers to the sensitivity analysis: main effect for IAT condition.

## Model

```{r, message=FALSE, warning=FALSE}
model_2 <- lm(d ~ IAT_condition + modern_racism_scale_total,
              contrasts = list(IAT_condition = "contr.sum"),
              data = passers_df)
```

## Summary

```{r, message=FALSE, warning=FALSE, echo=FALSE}
sjt.lm(model_2, 
       emph.p = FALSE, 
       show.std = TRUE)
```

## Plot effects

```{r, message=FALSE, fig.height=4, fig.width=6, warning=FALSE, echo=FALSE}
sjp.lm(model_2, 
       type = "fe")
```

# Hypothesis test 3: differences in response bias (c)

H3 refers to the response bias analysis: main effect for IAT condition.

## Model

```{r, message=FALSE, warning=FALSE}
model_3 <- lm(c ~ IAT_condition + modern_racism_scale_total,
              contrasts = list(IAT_condition = "contr.sum"),
              data = passers_df)
```

## Summary

```{r, message=FALSE, warning=FALSE, echo=FALSE}
sjt.lm(model_3, 
       emph.p = FALSE, 
       show.std = TRUE)
```

## Plot effects

```{r, message=FALSE, fig.height=4, fig.width=6, warning=FALSE, echo=FALSE}
sjp.lm(model_3, 
       type = "fe")
```

# Hypothesis test 4: Differences in self-reported ratings between contrast pair conditions

H4: Participants who complete a race IAT will demonstrate more negative self-reported evaluations of the outgroup (black people) on the ratings than participants who completed a control (flowers-insects) IAT.  

```{r, message=FALSE, warning=FALSE, echo=FALSE}
ratings_data <- 
  read.csv("data/processed/processed ratings data - long format.csv") %>%
  dplyr::mutate(participant = as.factor(participant)) %>%
  filter(IAT_exclude_based_on_fast_trials == FALSE & 
           shooter_exclude_based_on_fast_trials == FALSE)
```

## Linear mixed model

### Model 

```{r, message=FALSE, warning=FALSE}
model_4 <- lmer(rating ~ IAT_condition + modern_racism_scale_total + (1 | participant), 
                contrasts = list(IAT_condition = "contr.sum"),  
                data = ratings_data)
```

### Summary

```{r, message=FALSE, warning=FALSE, echo=FALSE}
sjt.lmer(model_4, 
         emph.p = FALSE, 
         p.kr = FALSE,
         show.std = TRUE)
```

### Plot effects

```{r, echo=FALSE, fig.height=3, fig.width=5, message=FALSE, warning=FALSE}
sjp.lmer(model_4, 
         p.kr = FALSE,
         type = "fe")
```

### Table predictions

```{r, message=FALSE, fig.height=3, fig.width=5, warning=FALSE, echo=FALSE}
model_4_predicted_effects <- as.data.frame(effect("IAT_condition", model_4))

model_4_predicted_effects %>% 
  round_df(2) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```
  
### Plot predictions

```{r, message=FALSE, fig.height=3, fig.width=4, warning=FALSE, echo=FALSE}
ggplot(data = model_4_predicted_effects,
       aes(x = IAT_condition, y = fit)) +
  geom_pointrange(aes(ymax = upper,
                      ymin = lower)) +
  ylab("Rating") +
  xlab("IAT condition") +
  scale_colour_grey() +
  theme_classic() 
```
