---
title: "Changing established atttiudes via the IAT: Experiment 3"
author: "Ian Hussey"
params:
  location_of_exp_1_data: ~/git/Analogical learning via the IAT II/experiment 1/
  location_of_exp_2_data: ~/git/Analogical learning via the IAT II/experiment 2/
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(message=FALSE,
                      warning=FALSE,
                      cache.lazy=FALSE)
```

```{r}

# dependencies
library(tidyverse)
library(psych)
library(weights)  # for rd(), a round() alternative
library(plotrix)  # for std.error
library(lme4)
library(effects)
library(knitr)
library(kableExtra)
library(sjPlot)
library(kableExtra)
library(apa)

# comment out if knitting to pdf
options(knitr.table.format = "html")

# knitr options
knitr::opts_chunk$set(echo = TRUE)

# rounds all numerics in a df
round_df <- function(df, digits) {
  nums <- vapply(df, is.numeric, FUN.VALUE = logical(1))
  df[,nums] <- round(df[,nums], digits = digits)
  (df)
}

```

# Descriptive statistics

## Full sample

```{r}

# get data
data_df <- 
  read.csv("data/processed/processed summary data - wide format.csv") %>%
  dplyr::mutate(gender = as.factor(gender))

#colnames(data_df)

attrition_rate <- data_df %>%
  dplyr::summarize(rate = round(sum(exclude)/n()*100, 1)) %>%
  as.numeric()

chi2_result <- data_df %>%
  count(IAT_condition, exclude) %>%
  spread(exclude, n) %>%
  select(-IAT_condition) %>%
  chisq.test(.) %>%
  chisq_apa(., print_n = FALSE, format = c("rmarkdown"), info = FALSE, print = FALSE)

```

Attrition rate = `r attrition_rate`%.

A chi squared test was used to assess whether the conditions suffered from differential attrition rates: `r chi2_result`.

## Analytic sample

I.e., after applying inclusion/exclusion criteria.

### Age

```{r}

passers_df <- data_df %>%
  filter(exclude == FALSE) 

passers_df %>% 
  dplyr::select(age) %>%
  psych::describe(fast = TRUE,  # subset of descriptive stats
                  ranges = FALSE,
                  trim = 0) %>%
  dplyr::select(-vars, -se) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

### Gender

```{r}

passers_df %>% 
  dplyr::count(gender) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

### Conditions

```{r}

passers_df %>% 
  dplyr::count(IAT_condition) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

### Performance on the IAT and shooter task

```{r}

passers_df %>% 
  dplyr::select(IAT_mean_RT,
                IAT_perc_acc,
                shooter_mean_RT, 
                shooter_perc_acc) %>%
  psych::describe(fast = TRUE,  # subset of descriptive stats
                  ranges = FALSE,
                  trim = 0) %>%
  dplyr::select(-vars, -se) %>%
  mutate(mean = round(mean, 2),
         sd = round(sd, 2)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

# Hypothesis test 1: Differences in Shooter task RT effects between IAT conditions

H1-H3: Completing an IAT serves to train attitudes as well as test them. Participants who complete a race IAT will influence implicit biases towards black people on the on the shooter task relative to participants who completed a control (flowers-insects) IAT.

H1 refers to the RT analysis: interaction effect between condition and trial type. Preregistered code mistakenly referred to the main effect too, but a difference in responding speed between conditions does not represent a racial bias. 

## Outlier removal

Exclude data points > 2.5 SD from mean.

```{r}

# exclude participants who had more than 10% of trials in less than 300ms, and make variables factors
shooter_data <- 
  read.csv("data/processed/processed shooter task data - long format.csv") %>%
  dplyr::mutate(participant = as.factor(participant),
                trial_type = as.factor(trial_type),
                IAT_condition = as.factor(IAT_condition)) %>%
  filter(IAT_exclude_based_on_fast_trials == FALSE)

# exclude outlier rts
shooter_data_outliers_removed <- shooter_data %>%
  schoRsch::outlier(dv = "rt",
                    todo="elim", 
                    upper.z = 2.5, 
                    lower.z = -2.5)

```

```{r, include=FALSE}

# Check that variables that should be factors are indeed factors
sapply(shooter_data_outliers_removed, class)

```

### Plot distribution

Distribution of shooter RTs after trimming outliers.

```{r fig.height=3, fig.width=5}

ggplot(data = shooter_data_outliers_removed) + 
  geom_density(aes(x = rt)) +
  theme_classic()

```

## Linear mixed model

### Model

```{r}

model_1 <- lmer(rt ~ trial_type * IAT_condition + modern_racism_scale_total + (1 | participant), 
                contrasts = list(trial_type = "contr.sum", IAT_condition = "contr.sum"),  
                data = shooter_data_outliers_removed)

```

- Hypothesis refers to the main effect for IAT condition or the IAT condition*trial type interaction effect.
- Confidence intervals via Wald estimation.

### Summary

```{r message=FALSE, warning=FALSE}

sjt.lmer(model_1, 
         emph.p = FALSE, 
         p.kr = FALSE,
         show.std = TRUE,
         digits.est = 3,
         digits.ci = 3)

```

### Plot effects

```{r fig.height=3, fig.width=5}

sjp.lmer(model_1, 
         p.kr = FALSE,
         type = "fe")

```

### Table predictions

```{r fig.height=3, fig.width=5}

model_1_predicted_effects <- 
  as.data.frame(effect("trial_type:IAT_condition", model_1)) %>%
  dplyr::rename(`IAT condition` = IAT_condition,
         `Trial type` = trial_type)

model_1_predicted_effects %>% 
  round_df(2) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```
  
### Plot predictions

```{r fig.height=3, fig.width=5}

ggplot(data = model_1_predicted_effects,
       aes(x = `Trial type`, y = fit, colour = `IAT condition`)) +
  geom_pointrange(aes(ymax = upper,
                      ymin = lower),
                  position = position_dodge(width = .1)) +
  geom_line(aes(group = `IAT condition`),
            position = position_dodge(width = .1)) +
  ylab("RT") +
  xlab("Trial type") +
  scale_colour_grey() +
  theme_classic()
  # theme(legend.justification=c(0.5,1), legend.position=c(0.5,1),
  #       legend.key = element_rect(colour = "transparent"))

```

# Hypothesis test 2: differences in Shooter task sensitivity (d') between IAT conditions

H2 refers to the sensitivity analysis: main effect for IAT condition.

## Model

```{r}

model_2 <- lm(d ~ IAT_condition + modern_racism_scale_total,
              contrasts = list(IAT_condition = "contr.sum"),
              data = passers_df)

```

## Summary

```{r}

sjt.lm(model_2, 
       emph.p = FALSE, 
       show.std = TRUE,
       digits.est = 2,
       digits.ci = 2)

```

## Plot effects

```{r fig.height=4, fig.width=6}

sjp.lm(model_2, 
       type = "fe")

```

# Hypothesis test 3: differences in Shooter task response bias (c) between conditions

H3 refers to the response bias analysis: main effect for IAT condition.

## Model

```{r}

model_3 <- lm(c ~ IAT_condition + modern_racism_scale_total,
              contrasts = list(IAT_condition = "contr.sum"),
              data = passers_df)

```

## Summary

```{r}

sjt.lm(model_3, 
       emph.p = FALSE, 
       show.std = TRUE,
       digits.est = 3,
       digits.ci = 3)

```

## Plot effects

```{r fig.height=4, fig.width=6}

sjp.lm(model_3, 
       type = "fe")

```

# Hypothesis test 4: Differences in self-reported ratings between IAT conditions

H4: Participants who complete a race IAT will demonstrate more negative self-reported evaluations of the outgroup (black people) on the ratings than participants who completed a control (flowers-insects) IAT.  

```{r}

ratings_data <- 
  read.csv("data/processed/processed ratings data - long format.csv") %>%
  dplyr::mutate(participant = as.factor(participant)) %>%
  mutate(IAT_condition = fct_relevel(as.factor(IAT_condition), "Race IAT", "Flowers-Insects IAT")) %>%
  filter(IAT_exclude_based_on_fast_trials == FALSE)

```

## Linear mixed model

### Model 

```{r}

model_4 <- lmer(rating ~ IAT_condition + modern_racism_scale_total + (1 | participant), 
                contrasts = list(IAT_condition = "contr.sum"),  
                data = ratings_data)

```

### Summary

```{r}

sjt.lmer(model_4, 
         emph.p = FALSE, 
         p.kr = FALSE,
         show.std = TRUE)

```

### Plot effects

```{r fig.height=3, fig.width=5}

sjp.lmer(model_4, 
         p.kr = FALSE,
         type = "fe")

```

### Table predictions

```{r fig.height=3, fig.width=5}

model_4_predicted_effects <- as.data.frame(effect("IAT_condition", model_4))

model_4_predicted_effects %>% 
  round_df(2) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```
  
### Plot predictions

```{r fig.height=3, fig.width=4}

ggplot(data = model_4_predicted_effects,
       aes(x = IAT_condition, y = fit)) +
  geom_pointrange(aes(ymax = upper,
                      ymin = lower)) +
  ylab("Rating") +
  xlab("IAT condition") +
  scale_colour_grey() +
  theme_classic() 

```

# Meta-analysis of self-report effects

## Descriptive statistics

```{r}

# get data from exp 1
setwd(params$location_of_exp_1_data)

ratings_data_exp1 <- 
  read.csv("data/processed/processed ratings data - long format.csv") %>%
  dplyr::mutate(experiment = 1,
                unique_id = as.factor(paste(experiment, participant, sep = "_")),
                IAT_condition = as.factor(IAT_condition),
                exclude = ifelse(IAT_exclude_based_on_fast_trials == TRUE, TRUE, 
                                 ifelse(SCIAT_exclude_based_on_fast_trials == TRUE, TRUE,
                                        FALSE))) %>%
  dplyr::select(unique_id, experiment, IAT_condition, exclude, rating, modern_racism_scale_total)


# get data from exp 2
setwd(params$location_of_exp_2_data)

ratings_data_exp2 <- 
  read.csv("data/processed/processed ratings data - long format.csv") %>%
  dplyr::mutate(experiment = 2,
                unique_id = as.factor(paste(experiment, participant, sep = "_")),
                IAT_condition = as.factor(IAT_condition),
                exclude = ifelse(IAT_exclude_based_on_fast_trials == TRUE, TRUE, FALSE)) %>%
  dplyr::select(unique_id, experiment, IAT_condition, exclude, rating, modern_racism_scale_total)

# trim exp 3 data
ratings_data_exp3 <-
  ratings_data %>%
  dplyr::mutate(experiment = 3,
                unique_id = as.factor(paste(experiment, participant, sep = "_")),
                exclude = ifelse(IAT_exclude_based_on_fast_trials == TRUE, TRUE, FALSE)) %>%
  dplyr::select(unique_id, experiment, IAT_condition, exclude, rating, modern_racism_scale_total)

# combine
combined_ratings_data <- rbind(ratings_data_exp1, ratings_data_exp2, ratings_data_exp3) %>%
  filter(exclude == FALSE) %>%
  mutate(IAT_condition = fct_relevel(as.factor(IAT_condition), "Race IAT", "Flowers-Insects IAT"))

```

## Linear mixed model

### Model 

```{r}

# http://www.metafor-project.org/doku.php/tips:rma_vs_lm_lme_lmer
model_meta <- lmer(rating ~ IAT_condition + modern_racism_scale_total + (1 | experiment) + (1 | unique_id),
                   contrasts = list(IAT_condition = "contr.sum"),  
                   data = combined_ratings_data)

```

- Hypothesis refers to main effect for IAT condition.
- *p* values estimated via Wald method.

### Summary

```{r}

sjt.lmer(model_meta, 
         emph.p = FALSE, 
         p.kr = FALSE,
         show.std = TRUE)

```

total N for meta analysis = `r combined_ratings_data %>% distinct(unique_id) %>% dplyr::summarize(n = n()) %>% as.numeric()`.

### Plot effects

```{r fig.height=3, fig.width=5}

sjp.lmer(model_meta, 
         p.kr = FALSE,
         type = "fe")

```

### Table predictions

```{r fig.height=3, fig.width=5}

model_meta_predicted_effects <- as.data.frame(effect("IAT_condition", model_meta))

model_meta_predicted_effects %>% 
  round_df(2) %>%
  kable()

```
  
### Plot predictions

```{r fig.height=2, fig.width=3}

ggplot(data = model_meta_predicted_effects,
       aes(x = IAT_condition, y = fit)) +
  geom_pointrange(aes(ymax = upper,
                      ymin = lower)) +
  ylab("Rating") +
  xlab("IAT condition") +
  scale_colour_grey() +
  theme_classic()

```