---
title: "Changing established atttiudes via the IAT: Experiment 2"
author: "Ian Hussey"
params:
  location_of_exp_1_data: ~/git/Analogical learning via the IAT II/experiment 1/
output:
  html_document:
    highlight: haddock
    toc: yes
---

Conceptual replication of Experiment 1. Employs AMP as DV rather than SC-IAT.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# dependencies
library(tidyverse)
library(psych)
library(weights)  # for rd(), a round() alternative
library(plotrix)  # for std.error
library(lme4)
library(effects)
library(knitr)
library(sjPlot)

# rounds all numerics in a df
round_df <- function(df, digits) {
  nums <- vapply(df, is.numeric, FUN.VALUE = logical(1))
  df[,nums] <- round(df[,nums], digits = digits)
  (df)
}
```

# Descriptive statistics

## Full sample

### Age

```{r, message=FALSE, warning=FALSE, echo = FALSE}
# get data
data_df <- 
  read.csv("data/processed/processed summary data - wide format.csv") %>%
  mutate(gender = as.factor(gender))

#colnames(data_df)

data_df %>% 
  dplyr::select(age) %>%
  psych::describe(fast = TRUE,  # subset of descriptive stats
                  ranges = FALSE,
                  trim = 0) %>%
  dplyr::select(-vars, -se) %>%
  kable()
```

### Gender

```{r, message=FALSE, warning=FALSE, echo = FALSE}
data_df %>% 
  dplyr::count(gender) %>%
  kable()
```

## Analytic sample

I.e., after applying inclusion/exclusion criteria.

### Conditions

```{r, message=FALSE, warning=FALSE, echo=FALSE}
passers_df <- 
  data_df %>%
  filter(IAT_exclude_based_on_fast_trials == FALSE & 
           exclude == FALSE) 

passers_df %>% 
  dplyr::count(IAT_condition) %>%
  kable()
```

### Performance on the IAT and SC-IAT

```{r, message=FALSE, warning=FALSE, echo=FALSE}
passers_df %>% 
  dplyr::select(IAT_mean_RT,
                IAT_perc_acc,
                AMP_mean_RT,
                AMP_perc_acc) %>%
  psych::describe(fast = TRUE,  # subset of descriptive stats
                  ranges = FALSE,
                  trim = 0) %>%
  dplyr::select(-vars, -se) %>%
  mutate(mean = round(mean, 2),
         sd = round(sd, 2)) %>%
  kable()
```

# Hypothesis test 1: Differences in IAT effects between contrast pair conditions

H1: Completing an IAT serves to train attitudes as well as test them. Participants who complete a race IAT will demonstrate more negative implicit evaluations of the outgroup (black people) on the AMP than participants who completed a control (flowers-insects) IAT.  

```{r, message=FALSE, warning=FALSE, include=FALSE}
AMP_data <- 
  read.csv("data/processed/processed AMP data - long format.csv") %>%
  dplyr::filter(exclude == FALSE,
                IAT_exclude_based_on_fast_trials == FALSE) %>%
  dplyr::mutate(participant = as.factor(participant),
                rating_factor = as.factor(ifelse(accuracy == 0, "negative", 
                                                 ifelse(accuracy == 1, "positive", NA))),
                IAT_condition = as.factor(IAT_condition)) %>%
  dplyr::rename(prime_type = item,
                rating = accuracy) 

# Check that variables that should be factors are indeed factors
sapply(AMP_data, class)
```

### Plot distribution

Distribution of AMP ratings.

```{r, echo=FALSE, fig.height=3, fig.width=5, message=FALSE, warning=FALSE}
AMP_participant_summary_data <-
  AMP_data %>%
  group_by(participant) %>%
  dplyr::summarize(perc_rating = round(mean(rating), 2))

ggplot(AMP_participant_summary_data, 
       aes(perc_rating)) + 
  geom_density(alpha=0.50) +
  theme_classic()
```

## Generalised linear mixed model

### Model

```{r, message=FALSE, warning=FALSE}
model_1 <- glmer(rating ~ prime_type * IAT_condition + modern_racism_scale_total + (1 | participant), 
                 family = "binomial",
                 contrasts = list(prime_type = "contr.sum", IAT_condition = "contr.sum"),  
                 data = AMP_data)
```

- Hypothesis refers to the IAT_condition*AMP_prime_type interaction effect.

### Summary

```{r, echo=FALSE, message=FALSE, warning=FALSE}
sjt.glmer(model_1, 
          emph.p = FALSE,
          digits.p = 7)

# precise p value
temp <- summary(model_1)$coefficients %>% as.tibble()
precise_interaction_p_value <- temp$`Pr(>|z|)`[5]
```

NB Precise *p* value for interaction = `r precise_interaction_p_value`.

### Plot effects

```{r, echo=FALSE, fig.height=3, fig.width=5, message=FALSE, warning=FALSE}
sjp.glmer(model_1, 
          type = "fe")
```

### Table predictions

```{r, echo=FALSE, fig.height=3, fig.width=5, message=FALSE, warning=FALSE}
model_1_predicted_effects <- 
  as.data.frame(effect("prime_type:IAT_condition", model_1)) %>%
  rename(`IAT condition` = IAT_condition,
         `AMP prime type` = prime_type)

model_1_predicted_effects %>% 
  round_df(2) %>%
  kable()
```
  
### Plot predictions

```{r, echo=FALSE, fig.height=3, fig.width=5, message=FALSE, warning=FALSE}
ggplot(data = model_1_predicted_effects,
       aes(x = `AMP prime type`, y = fit, colour = `IAT condition`)) +
  geom_pointrange(aes(ymax = upper,
                      ymin = lower),
                  position = position_dodge(width = .1)) +
  geom_line(aes(group = `IAT condition`),
            position = position_dodge(width = .1)) +
  ylab("Mean rating of Chinese characters") +
  xlab("AMP prime type") +
  scale_colour_grey() +
  theme_classic() 
```

# Hypothesis test 2: Differences in self-reported ratings between contrast pair conditions

H2: Completing an IAT serves to train attitudes as well as test them. Participants who complete a race IAT will demonstrate more negative self-reported evaluations of the outgroup (black people) on the ratings than participants who completed a control (flowers-insects) IAT.  

```{r, message=FALSE, warning=FALSE, echo=FALSE}
ratings_data <- 
  read.csv("data/processed/processed ratings data - long format.csv") %>%
  mutate(participant = as.factor(participant)) %>%
  filter(IAT_exclude_based_on_fast_trials == FALSE & 
           exclude == FALSE)
```

## Linear mixed model

### Model 

```{r, message=FALSE, warning=FALSE}
model_2 <- lmer(rating ~ IAT_condition + modern_racism_scale_total + (1 | participant), 
                contrasts = list(IAT_condition = "contr.sum"),  
                data = ratings_data)
```

- Hypothesis refers to main effect for IAT condition.
- *p* values estimated via Wald method.

### Summary

```{r, message=FALSE, warning=FALSE, echo=FALSE}
sjt.lmer(model_2, 
         emph.p = FALSE, 
         p.kr = FALSE,
         show.std = TRUE)
```

### Plot effects

```{r, echo=FALSE, fig.height=3, fig.width=5, message=FALSE, warning=FALSE}
sjp.lmer(model_2, 
         p.kr = FALSE,
         type = "fe")
```

### Table predictions

```{r, message=FALSE, fig.height=3, fig.width=5, warning=FALSE, echo=FALSE}
model_2_predicted_effects <- as.data.frame(effect("IAT_condition", model_2))

model_2_predicted_effects %>% 
  round_df(2) %>%
  kable()
```
  
### Plot predictions

```{r, message=FALSE, fig.height=3, fig.width=4, warning=FALSE, echo=FALSE}
ggplot(data = model_2_predicted_effects,
       aes(x = IAT_condition, y = fit)) +
  geom_pointrange(aes(ymax = upper,
                      ymin = lower)) +
  ylab("Rating") +
  xlab("IAT condition") +
  scale_colour_grey() +
  theme_classic()
```

# Hypothesis test 3: Combination analysis of self-report data

H3: H2 will also be applied to a combination analysis of all self-report ratings data collected in both Experiments 1 and 2.

## Descriptive statistics

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# trim exp 2 data
ratings_data_exp2 <-
  ratings_data %>%
  dplyr::mutate(experiment = 2,
                unique_id = as.factor(paste(experiment, participant, sep = "_"))) %>%
  dplyr::select(unique_id, experiment, IAT_condition, exclude, rating, modern_racism_scale_total)

# get data from exp 1
setwd(params$location_of_exp_1_data)

ratings_data_exp1 <- 
  read.csv("data/processed/processed ratings data - long format.csv") %>%
  #dplyr::filter(exclude == FALSE) %>%
  dplyr::mutate(experiment = 1,
                unique_id = as.factor(paste(experiment, participant, sep = "_")),
                IAT_condition = as.factor(IAT_condition),
                exclude = ifelse(IAT_exclude_based_on_fast_trials == TRUE, TRUE, 
                                 ifelse(SCIAT_exclude_based_on_fast_trials == TRUE, TRUE,
                                        FALSE))) %>%
  dplyr::select(unique_id, experiment, IAT_condition, exclude, rating, modern_racism_scale_total)

combined_ratings_data <- rbind(ratings_data_exp1, ratings_data_exp2)
```

## Linear mixed model

### Model 

```{r, message=FALSE, warning=FALSE}
model_3 <- lmer(rating ~ IAT_condition + modern_racism_scale_total + (1 | experiment) + (1 | unique_id), 
                contrasts = list(IAT_condition = "contr.sum"),  
                data = combined_ratings_data)
```

- Hypothesis refers to main effect for IAT condition.
- *p* values estimated via Wald method.

### Summary

```{r, message=FALSE, warning=FALSE, echo=FALSE}
sjt.lmer(model_3, 
         emph.p = FALSE, 
         p.kr = FALSE,
         show.std = TRUE)
```

### Plot effects

```{r, echo=FALSE, fig.height=3, fig.width=5, message=FALSE, warning=FALSE}
sjp.lmer(model_3, 
         p.kr = FALSE,
         type = "fe")
```

### Table predictions

```{r, message=FALSE, fig.height=3, fig.width=5, warning=FALSE, echo=FALSE}
model_3_predicted_effects <- as.data.frame(effect("IAT_condition", model_3))

model_3_predicted_effects %>% 
  round_df(2) %>%
  kable()
```
  
### Plot predictions

```{r, message=FALSE, fig.height=3, fig.width=4, warning=FALSE, echo=FALSE}
ggplot(data = model_3_predicted_effects,
       aes(x = IAT_condition, y = fit)) +
  geom_pointrange(aes(ymax = upper,
                      ymin = lower)) +
  ylab("Rating") +
  xlab("IAT condition") +
  scale_colour_grey() +
  theme_classic()
```
