---
title: "Changing established atttiudes via the IAT: Experiment 1"
author: "Ian Hussey"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r}
knitr::opts_chunk$set(message=FALSE,
                      warning=FALSE,
                      cache.lazy=FALSE)
```

```{r}

# dependencies
library(tidyverse)
library(psych)
library(weights)  # for rd(), a round() alternative
library(plotrix)  # for std.error
library(lme4)
library(effects)
library(knitr)
library(kableExtra)
library(sjPlot)
library(apa)

# rounds all numerics in a df
round_df <- function(df, digits) {
  nums <- vapply(df, is.numeric, FUN.VALUE = logical(1))
  df[,nums] <- round(df[,nums], digits = digits)
  (df)
}

# options
options(knitr.table.format = "html") # necessary configuration of tables

# disable scientific notation
options(scipen = 999) 

```

# Descriptive statistics

## Full sample

Assess selective attrition between conditions.

```{r}

# get data
data_df <- 
  read.csv("data/processed/processed summary data - wide format.csv") %>%
  dplyr::mutate(gender = as.factor(gender))

#colnames(data_df)

attrition_rate <- data_df %>%
  dplyr::summarize(rate = round(sum(exclude)/n()*100, 1)) %>%
  as.numeric()

data_df %>%
  count(exclude) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

chi2_result <- data_df %>%
  count(IAT_condition, exclude) %>%
  spread(exclude, n) %>%
  select(-IAT_condition) %>%
  chisq.test(.) %>%
  chisq_apa(., print_n = TRUE, format = c("rmarkdown"), info = FALSE, print = FALSE)

```

Attrition rate = `r attrition_rate`%.

A chi squared test was used to assess whether the conditions suffered from differential attrition rates: `r chi2_result`. [nb results of p = 1.0 will show as = .000 due to an error in how chisq_apa trims its results]

## Analytic sample

I.e., after applying inclusion/exclusion criteria.

### Age

```{r}

passers_df <- data_df %>%
  filter(exclude == FALSE) 

passers_df %>% 
  dplyr::select(age) %>%
  psych::describe(fast = TRUE,  # subset of descriptive stats
                  ranges = FALSE,
                  trim = 0) %>%
  dplyr::select(-vars, -se) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

### Gender

```{r}

passers_df %>% 
  dplyr::count(gender) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

### Performance on the IAT and SC-IAT

```{r}

passers_df %>% 
  dplyr::select(IAT_mean_RT,
                IAT_perc_acc,
                SCIAT_mean_RT,
                SCIAT_perc_acc) %>%
  psych::describe(fast = TRUE,  # subset of descriptive stats
                  ranges = FALSE,
                  trim = 0) %>%
  dplyr::select(-vars, -se) %>%
  dplyr::mutate(mean = round(mean, 2),
         sd = round(sd, 2)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

# Hypothesis test 1: Differences in IAT effects between contrast pair conditions

H1: Completing an IAT serves to train attitudes as well as test them. Participants who complete a race IAT will demonstrate more negative implicit evaluations of the outgroup (black people) on the SCIAT than participants who completed a control (flowers-insects) IAT.  

## Outlier removal

Exclude data points > 2.5 SD from mean.

```{r}

# exclude participants who had more than 10% of trials in less than 300ms, and make variables factors
SCIAT_data <- 
  read.csv("data/processed/processed SCIAT data - long format.csv") %>%
  dplyr::mutate(participant = as.factor(participant),
         block = as.factor(ifelse(block == "compatible", "black-positive", 
                        ifelse(block == "incompatible", "black-negative", NA))),
         IAT_condition = as.factor(IAT_condition)) %>%
  dplyr::rename(SCIAT_block = block) %>%
  filter(IAT_exclude_based_on_fast_trials == FALSE & 
           SCIAT_exclude_based_on_fast_trials == FALSE)  

# exclude outlier rts
SCIAT_data_outliers_removed <-
  SCIAT_data %>%
  schoRsch::outlier(dv = "rt",
                    todo = "elim", 
                    upper.z = 2.5, 
                    lower.z = -2.5)

```

```{r}

# Check that variables that should be factors are indeed factors
sapply(SCIAT_data_outliers_removed, class)

```

### Plot distribution

Distribution of SC-IAT RTs after trimming outliers.

```{r fig.height=3, fig.width=5}

ggplot(data = SCIAT_data_outliers_removed) + 
  geom_density(aes(x = rt)) +
  theme_classic()

```

## Linear mixed model

### Model

```{r}

model_1 <- lmer(rt ~ SCIAT_block * IAT_condition + modern_racism_scale_total + (1 | participant), 
                contrasts = list(SCIAT_block = "contr.sum", IAT_condition = "contr.sum"),  
                data = SCIAT_data_outliers_removed)

```

- Hypothesis refers to the IAT_condition*SCIAT_block interaction effect.
- Confidence intervals via Wald estimation.

### Summary

```{r}

sjt.lmer(model_1, 
         emph.p = FALSE, 
         p.kr = FALSE,
         show.std = TRUE,
         digits.est = 3,
         digits.ci = 3)

```

### Plot effects

```{r fig.height=3, fig.width=5}

sjp.lmer(model_1, 
         p.kr = FALSE,
         type = "fe")

```

### Table predictions

```{r fig.height=3, fig.width=5}

model_1_predicted_effects <- 
  as.data.frame(effect("SCIAT_block:IAT_condition", model_1)) %>%
  dplyr::rename(`IAT condition` = IAT_condition,
         `SCIAT block` = SCIAT_block)

model_1_predicted_effects %>% 
  round_df(2) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```
  
### Plot predictions

```{r fig.height=3, fig.width=5}

ggplot(data = model_1_predicted_effects,
       aes(x = `SCIAT block`, y = fit, colour = `IAT condition`)) +
  geom_pointrange(aes(ymax = upper,
                      ymin = lower),
                  position = position_dodge(width = .1)) +
  geom_line(aes(group = `IAT condition`),
            position = position_dodge(width = .1)) +
  ylab("RT") +
  xlab("SC-IAT block") +
  scale_colour_grey() +
  theme_classic()
  # theme(legend.justification=c(0.5,1), legend.position=c(0.5,1),
  #       legend.key = element_rect(colour = "transparent"))

```

## Alternative analytic strategy: Linear fixed effect model on SC-IAT D1 scores

For illustrative/reader familiarity purposes only.

### Model

```{r}

model_1_alt <- lm(SCIAT_D1 ~ IAT_condition + modern_racism_scale_total,
                  contrasts = list(IAT_condition = "contr.sum"),
                  data = passers_df)

```

### Summary

```{r}

sjt.lm(model_1_alt, 
       emph.p = FALSE, 
       show.std = TRUE)

```

### Plot effects

```{r fig.height=4, fig.width=6}

sjp.lm(model_1_alt, 
       type = "fe")

```

### Estimate effect size

Differences between groups are more meaningful when experessed as Cohen's d, so regress against racism (the covariate) and then compare group differences in the residuals.

```{r}

model_1a_alt <- lm(SCIAT_D1 ~ modern_racism_scale_total,
                   data = passers_df)

temp <- passers_df %>%
  rownames_to_column()

resid_data <- residuals(model_1a_alt) %>%
    as.data.frame()
colnames(resid_data) <- "residual_SCIAT_D1"
resid_data <- resid_data %>%  
  rownames_to_column() %>%
  left_join(temp, by = "rowname")

effsize::cohen.d(residual_SCIAT_D1 ~ IAT_condition, data = resid_data)

```

# Hypothesis test 2: Differences in self-reported ratings between contrast pair conditions

H2: Completing an IAT serves to train attitudes as well as test them. Participants who complete a race IAT will demonstrate more negative self-reported evaluations of the outgroup (black people) on the ratings than participants who completed a control (flowers-insects) IAT.  

```{r}

ratings_data <- 
  read.csv("data/processed/processed ratings data - long format.csv") %>%
  dplyr::mutate(participant = as.factor(participant)) %>%
  mutate(IAT_condition = fct_relevel(as.factor(IAT_condition), "Race IAT", "Flowers-Insects IAT")) %>%
  filter(IAT_exclude_based_on_fast_trials == FALSE & 
           SCIAT_exclude_based_on_fast_trials == FALSE)

```

## Linear mixed model

### Model 

```{r}

model_2 <- lmer(rating ~ IAT_condition + modern_racism_scale_total + (1 | participant), 
                contrasts = list(IAT_condition = "contr.sum"),  
                data = ratings_data)

```

### Summary

```{r}

sjt.lmer(model_2, 
         emph.p = FALSE, 
         p.kr = FALSE,
         show.std = TRUE)

```

### Plot effects

```{r fig.height=3, fig.width=5}

sjp.lmer(model_2, 
         p.kr = FALSE,
         type = "fe")

```

### Table predictions

```{r fig.height=3, fig.width=5}

model_2_predicted_effects <- as.data.frame(effect("IAT_condition", model_2))

model_2_predicted_effects %>% 
  round_df(2) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```
  
### Plot predictions

```{r fig.height=2, fig.width=2}

ggplot(data = model_2_predicted_effects,
       aes(x = IAT_condition, y = fit)) +
  geom_pointrange(aes(ymax = upper,
                      ymin = lower)) +
  ylab("Rating") +
  xlab("IAT condition") +
  scale_colour_grey() +
  theme_classic() 

```
