---
title: "Changing established atttiudes via the IAT: Experiment 4"
author: "Ian Hussey"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

Direct replication of Experiment 2.

```{r}
knitr::opts_chunk$set(message=FALSE,
                      warning=FALSE,
                      cache.lazy=FALSE)
```

```{r}

# dependencies
library(tidyverse)
library(psych)
library(weights)  # for rd(), a round() alternative
library(plotrix)  # for std.error
library(lme4)
library(effects)
library(knitr)
library(kableExtra)
library(sjPlot)
library(apa)

# rounds all numerics in a df
round_df <- function(df, digits) {
  nums <- vapply(df, is.numeric, FUN.VALUE = logical(1))
  df[,nums] <- round(df[,nums], digits = digits)
  (df)
}

# options
options(knitr.table.format = "html") # necessary configuration of tables

# disable scientific notation
options(scipen = 999) 

```

# Descriptive statistics

## Full sample

```{r}

# get data
data_df <- 
  read.csv("data/processed/processed summary data - wide format.csv") %>%
  mutate(gender = as.factor(gender))

#colnames(data_df)

attrition_rate <- data_df %>%
  dplyr::summarize(rate = round(sum(exclude)/n()*100, 1)) %>%
  as.numeric()

data_df %>%
  count(exclude) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

chi2_result <- data_df %>%
  count(IAT_condition, exclude) %>%
  spread(exclude, n) %>%
  select(-IAT_condition) %>%
  chisq.test(.) %>%
  chisq_apa(., print_n = TRUE, format = c("rmarkdown"), info = FALSE, print = FALSE)

passers_df <- data_df %>%
  filter(exclude == FALSE) 

N <- passers_df %>%
  dplyr::summarize(n = n()) %>%
  as.numeric()

```

Attrition rate = `r attrition_rate`%.

A chi squared test was used to assess whether the conditions suffered from differential attrition rates: `r chi2_result`. [nb results of p = 1.0 will show as = .000 due to an error in how chisq_apa trims its results]

## Analytic sample

I.e., after applying inclusion/exclusion criteria.

### Age

```{r}

passers_df %>% 
  dplyr::select(age) %>%
  psych::describe(fast = TRUE,  # subset of descriptive stats
                  ranges = FALSE,
                  trim = 0) %>%
  dplyr::select(-vars, -se) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

### Gender

```{r}

passers_df %>% 
  dplyr::count(gender) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

### Performance on the IAT and AMP

```{r}

passers_df %>% 
  dplyr::select(IAT_mean_RT,
                IAT_perc_acc,
                AMP_mean_RT,
                AMP_perc_acc) %>%
  psych::describe(fast = TRUE,  # subset of descriptive stats
                  ranges = FALSE,
                  trim = 0) %>%
  dplyr::select(-vars, -se) %>%
  mutate(mean = round(mean, 2),
         sd = round(sd, 2)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

# Hypothesis test 1: Differences in AMP effects between contrast pair conditions

H1: Completing an IAT serves to train attitudes as well as test them. Participants who complete a race IAT will demonstrate more negative implicit evaluations of the outgroup (black people) on the AMP than participants who completed a control (flowers-insects) IAT.  

```{r}

AMP_data <- 
  read.csv("data/processed/processed AMP data - long format.csv") %>%
  dplyr::filter(exclude == FALSE,
                IAT_exclude_based_on_fast_trials == FALSE) %>%
  dplyr::mutate(participant = as.factor(participant),
                rating_factor = as.factor(ifelse(accuracy == 0, "negative", 
                                                 ifelse(accuracy == 1, "positive", NA))),
                IAT_condition = as.factor(IAT_condition)) %>%
  dplyr::rename(prime_type = item,
                rating = accuracy)

# Check that variables that should be factors are indeed factors
sapply(AMP_data, class)

```

### Plot distribution

Distribution of AMP ratings.

```{r fig.height=3, fig.width=5}

AMP_participant_summary_data <-
  AMP_data %>%
  group_by(participant) %>%
  dplyr::summarize(perc_rating = round(mean(rating), 2))

ggplot(AMP_participant_summary_data, 
       aes(perc_rating)) + 
  geom_density(alpha=0.50) +
  theme_classic()

```

## Generalised linear mixed model

### Model

```{r}

model_1 <- glmer(rating ~ prime_type * IAT_condition + modern_racism_scale_total + (1 | participant), 
                 family = "binomial",
                 contrasts = list(prime_type = "contr.sum", 
                                  IAT_condition = "contr.sum"),  
                 data = AMP_data)

```

- Hypothesis refers to the IAT_condition*AMP_prime_type interaction effect.

### Summary

```{r message=FALSE, warning=FALSE}

sjt.glmer(model_1, 
          emph.p = FALSE,
          digits.p = 7)

```

### Plot effects

```{r fig.height=3, fig.width=5}

sjp.glmer(model_1, 
          type = "fe")

```

### Table predictions

```{r fig.height=3, fig.width=5}

model_1_predicted_effects <- 
  as.data.frame(effect("prime_type:IAT_condition", model_1)) %>%
  rename(`IAT condition` = IAT_condition,
         `AMP prime type` = prime_type)

model_1_predicted_effects %>% 
  round_df(2) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```
  
### Plot predictions

```{r fig.height=3, fig.width=5}

ggplot(data = model_1_predicted_effects,
       aes(x = `AMP prime type`, y = fit, colour = `IAT condition`)) +
  geom_pointrange(aes(ymax = upper,
                      ymin = lower),
                  position = position_dodge(width = .1)) +
  geom_line(aes(group = `IAT condition`),
            position = position_dodge(width = .1)) +
  ylab("Mean rating") +
  xlab("Prime type") +
  scale_colour_grey() +
  theme_classic() 

```

# Hypothesis test 2: Differences in self-reported ratings between contrast pair conditions

H2: Completing an IAT serves to train attitudes as well as test them. Participants who complete a race IAT will demonstrate more negative self-reported evaluations of the outgroup (black people) on the ratings than participants who completed a control (flowers-insects) IAT.  

```{r}

ratings_data <- 
  read.csv("data/processed/processed ratings data - long format.csv") %>%
  mutate(participant = as.factor(participant)) %>%
  mutate(IAT_condition = fct_relevel(as.factor(IAT_condition), "Race IAT", "Flowers-Insects IAT")) %>%
  filter(IAT_exclude_based_on_fast_trials == FALSE & 
           exclude == FALSE)

```

## Linear mixed model

### Model 

```{r}

model_2 <- lmer(rating ~ IAT_condition + modern_racism_scale_total + (1 | participant), 
                contrasts = list(IAT_condition = "contr.sum"),  
                data = ratings_data)

```

- Hypothesis refers to main effect for IAT condition.
- *p* values estimated via Wald method.

### Summary

```{r}

sjt.lmer(model_2, 
         emph.p = FALSE, 
         p.kr = FALSE,
         show.std = TRUE)

```

### Plot effects

```{r fig.height=3, fig.width=5}

sjp.lmer(model_2, 
         p.kr = FALSE,
         type = "fe")

```

### Table predictions

```{r fig.height=3, fig.width=5}

model_2_predicted_effects <- as.data.frame(effect("IAT_condition", model_2))

model_2_predicted_effects %>% 
  round_df(2) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```
  
### Plot predictions

```{r fig.height=2, fig.width=2}

ggplot(data = model_2_predicted_effects,
       aes(x = IAT_condition, y = fit)) +
  geom_pointrange(aes(ymax = upper,
                      ymin = lower)) +
  ylab("Rating") +
  xlab("IAT condition") +
  scale_colour_grey() +
  theme_classic()

```
